---
title: "Stroop analysis"
author: "Gary Lupyan"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: yes
    theme: cerulean
    code_folding: hide
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---


# Setup/Load libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(tidyverse)
library(ggplot2)
library(ggbeeswarm)
library(lme4)
theme_set(theme_bw(base_size = 12))
```


# Load data
```{r message=FALSE}
dat <- read.csv("../raw_data/stroopData.csv") %>% 
	mutate(RT = ifelse(RT<300,NA,RT),
				 RT_correct = ifelse(is_correct==0,NA,RT))
dat %>% summary
```

# Example of pivoting from long to wide

Long/Tidy format -- this is the format we'll be working with most of the time. One row per observation. Each column repersesents a specific variable.
```{r message=FALSE}
dat %>% 
	group_by(subj_code,orientation,trial_type) %>% 
	summarize(RT_correct = mean(RT_correct,na.rm=TRUE)) %>% head
```


Now let's pivot to wide which allows us to easily compute, e.g., by-subject differences for a condition, but loses the paradigm of one-column=one variable (notice that now RTs are split across multiple columns).

```{r message=FALSE}
dat %>% 
	group_by(subj_code,trial_type) %>% 
	summarize(RT_correct = mean(RT_correct,na.rm=TRUE)) %>% pivot_wider(names_from=c(trial_type),values_from=RT_correct) %>%
	mutate(congruence_effect = incongruent - congruent) %>% head
```


# Graphing

A basic interaction plot. Notice that we're first averaging the data and then passing the averages to ggplot.

```{r message=FALSE}
dat %>% 
	filter(prop_incongruent<=.5) %>%
	group_by(orientation,trial_type) %>% summarize(RT=mean(RT_correct,na.rm=TRUE)) %>%
	ggplot(aes(y=RT,x=trial_type,group=trial_type:orientation))+
	geom_point(aes(color=orientation),position = position_dodge(width = .1),size=4)+
  geom_line(aes(group=orientation),position = position_dodge(width = .1))+
	ylab("Reaction Time") + 
  xlab("Trial Type")+
	scale_color_brewer(palette = "Set1")

```

Here's a bar-chart version of the above graph. To my eye, the two main effects and their interaction is much harder to interpret, but your mileage may vary!

```{r message=FALSE}
dat %>% 
	filter(prop_incongruent<=.5) %>%
	group_by(orientation,trial_type) %>% summarize(RT=mean(RT_correct,na.rm=TRUE)) %>%
	ggplot(aes(y=RT,x=trial_type,group=trial_type:orientation))+
	geom_bar(aes(fill=orientation),position = position_dodge(width = .9), stat="identity")+
	ylab("Reaction Time") + 
  xlab("Trial Type") + 
	coord_cartesian(ylim = c(500, 900))+
  scale_fill_brewer(palette = "Set1")
```

An example of faceting by `prop_incongruent` to see how the orientation:trial_type interaction differs in the two .25 and .50 conditions

```{r message=FALSE}
dat %>% 
	filter(prop_incongruent<=.5 & is_correct==1) %>%
	group_by(prop_incongruent,orientation,trial_type) %>% summarize(RT=mean(RT)) %>%
	ggplot(aes(y=RT,x=trial_type,group=trial_type:orientation))+
	geom_point(aes(color=orientation),position = position_dodge(width = .1),size=4)+
  geom_line(aes(group=orientation),position = position_dodge(width = .1))+
	ylab("Reaction Time") + 
  xlab("Trial Type") + 
  scale_color_brewer(palette = "Set1")+
	facet_wrap(~prop_incongruent)
```
Let's combine a few graphing types to show RT distributions and individual RT means in each condition

```{r message=FALSE}
dat %>% 
	filter(prop_incongruent<=.5) %>%
	group_by(subj_code,trial_type) %>% summarize(RT=mean(RT_correct,na.rm=TRUE)) %>%
	ggplot(aes(y=RT,x=trial_type))+
  geom_violin() +
  ggbeeswarm::geom_beeswarm(aes(color=trial_type),alpha=.3,cex=.2,priority="density")+
  ylab("Reaction Time") + 
  xlab("Orientation") + 
	stat_summary(geom = "point", fun = mean, position = position_dodge(width = .1),shape=21, size=4, fill="red") +
	geom_line(aes(group=subj_code))+
  scale_color_brewer(palette = "Set1")

```


Same idea, but plotting accuracy instead
```{r message=FALSE}
dat %>% 
	filter(prop_incongruent<=.5) %>%
	group_by(subj_code,trial_type) %>% summarize(is_correct=mean(is_correct)) %>%
	ggplot(aes(y=is_correct,x=trial_type))+
  geom_violin() +
  geom_beeswarm(aes(color=trial_type),alpha=.3,cex=.2,priority="density")+
  ylab("Accuracy") + 
  xlab("Orientation") + 
	stat_summary(geom = "point", fun = mean, position = position_dodge(width = .1),shape=21, size=4, fill="red") +
	geom_line(aes(group=subj_code))+
	scale_color_brewer(palette = "Set1")

```



```{r message=FALSE}
dat %>% 
	filter(prop_incongruent<=.5 & is_correct==1) %>%
	group_by(subj_code,orientation,trial_type) %>% summarize(RT=mean(RT)) %>%
	ggplot(aes(y=RT,x=trial_type))+
  geom_violin() +
  geom_beeswarm(aes(color=trial_type),alpha=.3,cex=.2,priority="density")+
  ylab("Reaction Time") + 
  xlab("Orientation") + 
	stat_summary(geom = "point", fun = mean, position = position_dodge(width = .1),shape=21, size=4, fill="red") +
	geom_line(aes(group=subj_code))+
	facet_grid(~ orientation)+
	scale_color_brewer(palette = "Set1")



```

# Models

Here's the stats model for those who're curious (not modeling item effects here because we're uninterested in generalizing to other ones, but one could of course include those and also see if certain color conflicts are especially confusing etc.)


Set contrasts to sum-coding (essentially centering the factors) so that the main effects are interpreted correctly in the presence of interactions
```{r message=FALSE}
contrasts(dat$trial_type) = contr.sum(2)
contrasts(dat$orientation) = contr.sum(2)

```


Linear mixed effects model
```{r message=FALSE}
dat %>% filter(prop_incongruent<=.5) %>%
	lmer(RT_correct~scale(prop_incongruent)*trial_type*orientation+(orientation+trial_type|subj_code),data=.) %>% summary
```

